---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
    cache = TRUE,
    collapse = TRUE,
    comment = "#>",
    fig.path = "README-files/"
)
```

[![Travis-CI Build Status](https://travis-ci.org/jjchern/fips.svg?branch=master)](https://travis-ci.org/jjchern/fips)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/jjchern/fips?branch=master&svg=true)](https://ci.appveyor.com/project/jjchern/fips)

# About {fips}

The R package {fips} makes it easier to merge geographic identifiers such as state FIPS, county FIPS, urban-rural codes, BEA region codes, and census region and division codes. 

For an overview of regions of the United States, see the wiki page for [List of regions of the United States](https://en.wikipedia.org/wiki/List_of_regions_of_the_United_States). 

The following datasets are available:

- `fips::fips`, `fips::state`, and `fips::lower48`: 
    - State-level FIPS codes. [(Source)](https://www.census.gov/geo/reference/ansi_statetables.html)
    - `fips::fips` contains 50 states, the District of Columbia, and the Outlying Areas of the United States.
    - `fips::state` contains 50 states and the District of Columbia.
    - `fips::lower48` contains the 48 Contiguous Continental States and the District of Columbia.

- `fips::county`, `fips::county_ipums_usa`
    - County-level FIPS codes.
    - `fips::county` contains 2010 county-level FIPS codes. [(Source)](https://www.census.gov/geo/reference/codes/cou.html)
    - `fips::county_ipums_usa` contains counties that are identified through the microdata maintained by IPUMS. [(Source)](https://usa.ipums.org/usa-action/variables/COUNTYFIP#description_section)
    
- `fips::nchs_urc`: 
    - NCHS Urban-Rural Classification Scheme for Counties. [(Source)](https://www.cdc.gov/nchs/data_access/urban_rural.htm)
    
- `fips::bea_region`:
    - BEA Region codes. [(Source)](https://www.bea.gov/regional/docs/regions.cfm)
    
- `fips::census_region_division`:
    - Census Bureau region and division codes. [(Source)](https://www2.census.gov/programs-surveys/popest/geographies/2015/state-geocodes-v2015.txt)
    
I might add other crosswalks for OMB standard federal regions, federal reserve districts, courts of appeals circuits, and Agricultural Research Service regions.
    
Similar implementation in Stata:

- [{statastates} by `@wschpero`](https://github.com/wschpero/statastates)

Similar R packages:

- [{zipcodes} by `@jacobkap`](https://github.com/jacobkap/zipcodes)
- [The dataset `US_states` in {EconData} by `@floswald`](https://github.com/floswald/EconData)

# Installation

You can install the development version of {fips} from Github with:

``` r
# install.packages("remotes")
remotes::install_github("jjchern/fips")
```

Or install the most recent released version of {fips} from Github with:

``` r
remotes::install_github("jjchern/fips@v0.0.5")
```

# Usage

## State-level FIPS code

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
fips::state

# fips::fips includes FIPS code for other outlying areas
fips::fips 
fips::fips %>% tail(10)

# fips::lower48 includes the 48 continental states and DC
fips::lower48
```

## 2010 FIPS code for counties

```{r}
fips::county
```

## Counties Identified in IPUMS USA (2005-forward)

```{r}
fips::county_ipums_usa %>% 
    select(countyfip, 
           `2000 5% & 1% unwt, acs 2005`,
           `acs 2006-2011`, 
           `2010 10%, acs 2012-onward`) %>% 
    na.omit() %>% 
    select(countyfip)
```

## Census Region and Division Codes

```{r}
fips::census_region_division
```

```{r census_regions, echo=FALSE, warning=FALSE, message=FALSE}
right_join(usmapdata::state, fips::census_region_division,
           by = c("id" = "fips")) %>% 
    haven::as_factor() -> census_region_division

ggplot() +
    geom_map(data = census_region_division, map = census_region_division,
             aes(x = long, y = lat, map_id = id, fill = region_name),
             colour = alpha("white", 0.1), size = 0.2) +
    geom_map(data = usmapdata::state, map = usmapdata::state,
             aes(x = long, y = lat, map_id = id),
             colour = alpha("white", 0.75), fill = "NA") +        
    coord_map("albers", lat0 = 30, lat1 = 40) +
    viridis::scale_fill_viridis(option = "C", discrete = TRUE) +
    ggtitle("Component State for Census Regions") +
    ggthemes::theme_map() +
    theme(legend.position = c(.82, .4),
          legend.title = element_blank(), 
          plot.title = element_text(hjust = 0.5))
```

```{r census_division, echo=FALSE, warning=FALSE, message=FALSE}
ggplot() +
    geom_map(data = census_region_division, map = census_region_division,
             aes(x = long, y = lat, map_id = id, fill = division_name),
             colour = alpha("white", 0.1), size = 0.2) +
    geom_map(data = usmapdata::state, map = usmapdata::state,
             aes(x = long, y = lat, map_id = id),
             colour = alpha("white", 0.75), fill = "NA") +        
    coord_map("albers", lat0 = 30, lat1 = 40) +
    viridis::scale_fill_viridis(option = "C", discrete = TRUE) +
    ggtitle("Component State for Census Divisions") +
    ggthemes::theme_map() +
    theme(legend.position = c(.82, .2),
          legend.title = element_blank(), 
          plot.title = element_text(hjust = 0.5))
```

## BEA Region codes for states

```{r}
fips::bea_region
```

```{r bea_regions, echo=FALSE, warning=FALSE, message=FALSE}

right_join(usmapdata::state, fips::bea_region,
           by = c("id" = "fips")) %>% 
    haven::as_factor() -> bea_region

ggplot() +
    geom_map(data = bea_region, map = bea_region,
             aes(x = long, y = lat, map_id = id, fill = short_region_name),
             colour = alpha("white", 0.1), size = 0.2) +
    geom_map(data = usmapdata::state, map = usmapdata::state,
             aes(x = long, y = lat, map_id = id),
             colour = alpha("white", 0.75), fill = "NA") +        
    coord_map("albers", lat0 = 30, lat1 = 40) +
    viridis::scale_fill_viridis(option = "C", discrete = TRUE) +
    ggtitle("Component State for BEA Regions") +
    ggthemes::theme_map() +
    theme(legend.position = c(.82, .3),
          legend.title=element_blank(), 
          plot.title = element_text(hjust = 0.5))
```

## NCHS Urban Rural Codes

```{r}
fips::nchs_urc
```

```{r nchs_urc_2013, echo=FALSE, warning=FALSE, message=FALSE}

right_join(usmapdata::county, fips::nchs_urc,
           by = c("id" = "fips")) %>% 
    haven::as_factor() -> nchs_urc

ggplot() +
    geom_map(data = nchs_urc, map = nchs_urc,
             aes(x = long, y = lat, map_id = id, fill = code2013),
             colour = alpha("white", 0.1), size = 0.2) +
    geom_map(data = usmapdata::state, map = usmapdata::state,
             aes(x = long, y = lat, map_id = id),
             colour = alpha("white", 0.75), fill = "NA") +        
    coord_map("albers", lat0 = 30, lat1 = 40) +
    viridis::scale_fill_viridis(option = "B", discrete = TRUE) +
    ggtitle("2013 NCHS Urban-Rural Classification Scheme for Counties") +
    ggthemes::theme_map() +
    theme(legend.position = c(.82, .3),
          legend.title=element_blank(), 
          plot.title = element_text(hjust = 0.5))
```

```{r nchs_urc_2006, echo=FALSE, warning=FALSE, message=FALSE}

ggplot() +
    geom_map(data = nchs_urc, map = nchs_urc,
             aes(x = long, y = lat, map_id = id, fill = code2006),
             colour = alpha("white", 0.1), size = 0.2) +
    geom_map(data = usmapdata::state, map = usmapdata::state,
             aes(x = long, y = lat, map_id = id),
             colour = alpha("white", 0.75), fill = "NA") + 
    coord_map("albers", lat0 = 30, lat1 = 40) +
    viridis::scale_fill_viridis(option = "A", discrete = TRUE) +
    ggtitle("2006 NCHS Urban-Rural Classification Scheme for Counties") +
    ggthemes::theme_map() +
    theme(legend.position = c(.82, .3),
          legend.title=element_blank(), 
          plot.title = element_text(hjust = 0.5))
```

```{r nchs_urc_1990, echo=FALSE, warning=FALSE, message=FALSE}

nchs_urc %>% 
    filter(!is.na(code1990)) %>% 
    ggplot() +
    geom_map(map = nchs_urc,
             aes(x = long, y = lat, map_id = id, fill = code1990),
             colour = alpha("white", 0.1), size = 0.2) +
    geom_map(data = usmapdata::state, map = usmapdata::state,
             aes(x = long, y = lat, map_id = id),
             colour = alpha("white", 0.75), fill = "NA") +        
    coord_map("albers", lat0 = 30, lat1 = 40) +
    viridis::scale_fill_viridis(option = "C", discrete = TRUE) +
    ggtitle("1990 NCHS Urban-Rural Classification Scheme for Counties") +
    ggthemes::theme_map() +
    theme(legend.position = c(.82, .3),
          legend.title=element_blank(), 
          plot.title = element_text(hjust = 0.5))
```
